Near Optimal Behavior with Approximate State Abstraction

### 2016-02-03 paper/results_discussion.tex
merging results updates

%Explanation of what plots done
For each domain we report two quantities as a function of epsilon in Figures \ref{fig:main_empirical_results1} and \ref{fig:main_empirical_results2}. First, we compare the number of states in the abstract \ac{MDP} to different values of $\epsilon$. The smaller this value, the smaller the state space of the $\ac{MDP}$ that the agent must plan over. Second, we report the value under the abstract policy of the initial ground state. 50 Trials are gathered for each data point \enote{except ones where it doesn't matter (nChain and Minefield) -- is this worth noting?}

%Discuss results
Our empirical results corroborate our thesis -- that approximate state abstractions can form the basis of learnable and useful state aggregation criteria. In both NChain and Minefield, we observe that despite increasing $\epsilon$ to reduce the number of states which must be planned over, optimal behavior is either fully maintained (NChain) or optimal behavior is very nearly maintained (Minefield). Similarly for Taxi and $\epsilon$ between $.02$ and $.025$ we observe a reduction in the number of states that must be planned over while value is fully maintained. After $.025$, increased reduction in state space size comes at a cost of value. Lastly, as $\epsilon$ is increased Random smoothly reduces the number of abstract states which must be planned over with a corresponding cost in the value of the derived policy.

Our experimental results also demonstrate a noteworthy contrast between approximate state abstraction on goal-based \acp{MDP} and random \acp{MDP}. Taxi exhibits relative stability in state space size and behavior for epsilon up to $.02$ until both precipitously fall off. We attribute the sudden fall off in number of abstract states and value of the derived policy in Taxi to its goal-based nature; once certain information critical for optimal behavior is lost in the state aggregation, solving the goal -- and so acquiring any reward -- is rendered entirely unattainable. Conversely, in Random a great deal of near-optimal policies are available to the agent. Thus, even the information for optimal behavior is lost with an increase in $\epsilon$ there are a great deal of near-optimal policies available to the agent which are not lost in the state aggregation.

>>>>

%Explanation of what plots done
For each domain we report two quantities as a function of epsilon. First, we compare the number of states in the abstract \ac{MDP} to different values of $\epsilon$, shown in Figure~\ref{fig:main_empirical_results1}. The smaller this value, the smaller the state space of the $\ac{MDP}$ that the agent must plan over. Second, we report the value under the abstract policy of the initial ground state, shown in Figure~\ref{fig:main_empirical_results2}. 200 Trials are run for each data point.

%Discuss results
Our empirical results corroborate our thesis -- that approximate state abstractions can form the basis of learnable and useful state aggregation criteria. In both NChain and Minefield, we observe that despite increasing $\epsilon$ to reduce the number of states which must be planned over, optimal behavior is either fully maintained (NChain) or very nearly maintained (Minefield). Similarly for Taxi and $\epsilon$ between $.02$ and $.025$ we observe a reduction in the number of states in the Abstract MDP while value is fully maintained. After $.025$, increased reduction in state space size comes at a cost of value. Lastly, as $\epsilon$ is increased in the Random domain, there is a smooth reduction in the number of abstract states which must be planned over with a corresponding cost in the value of the derived policy.

Our experimental results also highlight a noteworthy characteristic of approximate state abstraction in goal-based \acp{MDP}. Taxi exhibits relative stability in state space size and behavior for epsilon up to $.02$ until both fall off dramatically. We attribute the sudden fall off in number of abstract states and value of the derived policy in Taxi to its goal-based nature; once information critical for achieving optimal behavior is lost in the state aggregation, solving the goal -- and so acquiring any reward -- is unattainable. Conversely, in the Random domain, a great deal of near-optimal policies are available to the agent. Thus, even the information for optimal behavior is lost with an increase in $\epsilon$ there are a great deal of near-optimal policies available to the agent which are not lost in the state aggregation.
